---
title: "Stat 331 Final Report"
output:
  html_document:
    code_folding: hide
---

#Andrew Yan, Joey Johnson, Karen Kauffman

###**Introduction:**
  
  For our final project we utilized a dataset that we found on kaggle on Pokemons from generation 1 to generation 6. The Pokemon data set comes in a clean easy to use format and it contains columns for each pokemon combat stat such as type, attack, defense and health. The dataset has a sizeable number of observations with 800 rows total. In our shiny app we aimed to find the differences in Pokemon stats between each generation and visualize these findings. We also wanted to see if the game is still well balanced and see if some types have better base stats than other types.
  
  
###**Usage:**

  Our application utilized tabs to navigate which widget you want to select. Within each widget contains a dropdown menu of for each item to select. The graphs and data tables related to each tab displays information on the dataset depending on what you selected on the dropdown menu.
  
  
###**Statistical Concepts:**

  In this project, we used confidence intervals to express "typical" stats for Pokemon of different types. Because the margin of error captures up to 2 standard deviations away from the mean, the bounds produced by the confidence intervals show that 99.7% of Pokemon of the given type should have the respective stat fall within those bounds. As a result, the interval is very large, but it seems appropriate - Pokemon games are designed to have some Pokemon who are easy to capture and relatively weak, and a few legendaries who are far harder to catch but have notably higher stats. Because there is such a large spectrum of Pokemon abilities, we expect a large standard deviation and in turn a large margin of error when calculating confidence intervals.
  
  
###**Plots to Illustrate Results:** 

*Distribution of Base Stats based on Generation. Displayed through a bar chart.*

From the data you can see that as you change Generations the general shape of the distribution changes.
For example if you were to take a look at the differences between the distributions of HP between Generations 1 and 6 you can see that there is only one peak for the Generation 1 histogram but the Generation 6 histogram has numerous peaks. Thus indicating that there is more variety for Generation 6 in the HP category.

In the shiny application the graph takes the user input for Generation and user input for which stat you want to create the bar chart for.
```{r, message = FALSE}
library(ggplot2)
library(fmsb)
library(tidyverse)
pokemon <- read.csv("Pokemon.csv")

ggplot(subset(pokemon, Generation == 1)) + 
      geom_histogram(aes_string("HP", fill = "Type.1")) +
      labs(title = str_c(c("Distribution of ", "HP", " for Gen ", 1), collapse = ""))

ggplot(subset(pokemon, Generation == 6)) + 
      geom_histogram(aes_string("HP", fill = "Type.1")) +
      labs(title = str_c(c("Distribution of ", "HP", " for Gen ", 6), collapse = ""))
```

*Confidence Intervals for average Combat Stats based on type. Displayed through a table.*

By using the confidence to find the average base combat stats for each generation we were able to
observe the difference in stats among each generation. Based on these confidence interavls it's easy to see that the game is still fairly balanced because each confidence interval is relatively similar to each other. Some variability within these confidence intervals can be due to the fact that some types have less observations. For example when finding the confidence intervals using "Fire" type as input there are only 52 observations as opposted to "Water" type pokemon with 112 observations.

In the shiny application it takes the user input for type and creates intervals from that input. Currently the input is set as the "Water" type.
```{r, message = FALSE} 
waterPokemon <- pokemon %>%
      filter(Type.1 == "Water")
    avgAttack<- mean(waterPokemon$Attack)
    avgDefense <- mean(waterPokemon$Defense)
    avgSPAtk <- mean(waterPokemon$Sp..Atk)
    avgSPDef <- mean(waterPokemon$Sp..Def)
    avgSpeed <- mean(waterPokemon$Speed)

    hpCI <- c(mean(waterPokemon$HP) - 2 * sd(waterPokemon$HP), mean(waterPokemon$HP) + 2 * sd(waterPokemon$HP))
    atkCI <- c(mean(waterPokemon$Attack) - 2 * sd(waterPokemon$Attack), mean(waterPokemon$Attack) + 2 * sd(waterPokemon$Attack))
    defCI <- c(mean(waterPokemon$Defense) - 2 * sd(waterPokemon$Defense), mean(waterPokemon$Defense) + 2 * sd(waterPokemon$Defense))
    spAtkCI <- c(mean(waterPokemon$Sp..Atk) - 2 * sd(waterPokemon$Sp..Atk), mean(waterPokemon$Sp..Atk) + 2 * sd(waterPokemon$Sp..Atk))
    spDefCI <- c(mean(waterPokemon$Sp..Def) - 2 * sd(waterPokemon$Sp..Def), mean(waterPokemon$Sp..Def) + 2 * sd(waterPokemon$Sp..Def))
    ci_df <- data_frame(Num_Observations = nrow(waterPokemon), Observation_Info = c("Lower Bound", "Upper Bound"), 
                        avg_HP_CI = hpCI, avg_ATK_CI = atkCI,
                        avg_DEF_CI = defCI, avg_SPATK_CI = spAtkCI, avg_SP_DEF_CI = spDefCI)
    ci_df[2, 1] <- ""
    ci_df
  
waterPokemon <- pokemon %>%
      filter(Type.1 == "Grass")
    avgAttack<- mean(waterPokemon$Attack)
    avgDefense <- mean(waterPokemon$Defense)
    avgSPAtk <- mean(waterPokemon$Sp..Atk)
    avgSPDef <- mean(waterPokemon$Sp..Def)
    avgSpeed <- mean(waterPokemon$Speed)

    hpCI <- c(mean(waterPokemon$HP) - 2 * sd(waterPokemon$HP), mean(waterPokemon$HP) + 2 * sd(waterPokemon$HP))
    atkCI <- c(mean(waterPokemon$Attack) - 2 * sd(waterPokemon$Attack), mean(waterPokemon$Attack) + 2 * sd(waterPokemon$Attack))
    defCI <- c(mean(waterPokemon$Defense) - 2 * sd(waterPokemon$Defense), mean(waterPokemon$Defense) + 2 * sd(waterPokemon$Defense))
    spAtkCI <- c(mean(waterPokemon$Sp..Atk) - 2 * sd(waterPokemon$Sp..Atk), mean(waterPokemon$Sp..Atk) + 2 * sd(waterPokemon$Sp..Atk))
    spDefCI <- c(mean(waterPokemon$Sp..Def) - 2 * sd(waterPokemon$Sp..Def), mean(waterPokemon$Sp..Def) + 2 * sd(waterPokemon$Sp..Def))
    ci_df <- data_frame(Num_Observations = nrow(waterPokemon), Observation_Info = c("Lower Bound", "Upper Bound"), 
                        avg_HP_CI = hpCI, avg_ATK_CI = atkCI,
                        avg_DEF_CI = defCI, avg_SPATK_CI = spAtkCI, avg_SP_DEF_CI = spDefCI)
    ci_df[2, 1] <- ""
    ci_df
```

###**Conclusion**

Overall, after playing with the data and analyzing a multitude a graphs we saw more variability for base stats among the different generations. In generation 1 typically the base stats were centered around one value and we saw more variety as we progress a few generations later. Of course this is not the case for every single stat like "Defense" as we saw that defense actually saw a dip in variability when comparing Generations 1 to 6.






